---
title: "Integrative_NMF"
output: html_document
date: '2022-06-29'
---

```{r setup, include=FALSE}
reticulate::use_condaenv("tensorflowenv2", required = TRUE)
library(knitr)
library(circlize)
library(tidyverse)
library(ComplexHeatmap)
library(viridis)
library(BiocStyle)
library(ButchR)
library(limma)
library(ggplot2)
library(riverplot)
library(ramify)
library(ggpubr)
```

# Loading data
Prepare NMF input: loading, normalization.

```{r}
set.seed(4563)

cell_names_filtered = c("GM12878", "HepG2", "IMR90")
cell_names_imputed = c('A549', 'HMEC', 'HUVEC', 'NHEK')

features <- c('H2AFZ', 'H3K27ac', 'H3K27me3', 'H3K36me3', 
              'H3K4me1', 'H3K4me2', 'H3K4me3', 'H3K79me2', 
              'H3K9ac', 'H3K9me3', 'H4K20me1')

paths1 <- lapply(cell_names_filtered, function(name){
  paste0("/media/ag-cherrmann/echernova/model_input/", name, "/positives_", name, "_filtered.csv")
})

paths2 <- lapply(cell_names_imputed, function(name){
  paste0("/media/ag-cherrmann/echernova/model_input/", name, "/positives_", name, "_imputed.csv")
})

paths <- append(paths1, paths2)

load_data <- function(path){
  
  data = read.csv(path)
  
  # Only get the central windows
  data = cbind(data[2:5], data[,grep('_0', names(data))])
  
  # Rename columns to be shorter
  new_names <- lapply(names(data[5:length(names(data))]), function(n){
    return(strsplit(n, '_')[[1]][2])
  })
  names(data) <- c(names(data[1:4]), new_names)
  
  # Return the input matrix for NMF
  matrix_V = data.matrix(data[5:length(names(data))])
  
  # Only keep the columns that correspond to the features vector
  # cell_name <- strsplit(new_names[[1]], '_')[[1]][1]
  # keep_cols <- lapply(features, function(n){
  #   return(paste0(cell_name, '_', n))
  # })
  matrix_V <- matrix_V[, features]
  
  # Delete NA
  matrix_V <- na.omit(matrix_V)
  
  # Normalize matrix
  matrix_V_norm <- normalizeQuantiles(matrix_V)
  
  return(matrix_V_norm)
}

matrices_V <- lapply(paths, load_data)
names(matrices_V)<-c(cell_names_filtered, cell_names_imputed)
```

```{r}
# Load TAD boundary conservation data
paths1_consv <- lapply(cell_names_filtered, function(name){
  paste0("/media/ag-cherrmann/echernova/tad_bound_conservation_10_40kb/", name, "_tad_bound_conservation_filtered.csv")
})

paths2_consv <- lapply(cell_names_imputed, function(name){
  paste0("/media/ag-cherrmann/echernova/tad_bound_conservation_10_40kb/", name, "_tad_bound_conservation_imputed.csv")
})

paths_consv <- append(paths1_consv, paths2_consv)

consv_list <- lapply(paths_consv, function(path){
  consv = read.csv(path)
  return(consv)
})
names(consv_list) <- c(cell_names_filtered, cell_names_imputed)
```


```{r}
# Load CTCF buinding score annotation
paths1_ctcf <- lapply(cell_names_filtered, function(name){
  paste0("/media/ag-cherrmann/echernova/CTCF_anno/", name, "_CTCF_filtered.csv")
})
# From imputed: only available for A549
paths2_ctcf <- "/media/ag-cherrmann/echernova/CTCF_anno/A549_CTCF_imputed.csv"

paths_ctcf <- c(paths1_ctcf, paths2_ctcf)
CTCF_anno <- lapply(paths_ctcf, function(path){
  return(read.csv(path, header = TRUE))
})
names(CTCF_anno) <- c(cell_names_filtered, "A549")
```


```{r}
# Load CTCF binary binding score annotation
paths1_ctcf <- lapply(cell_names_filtered, function(name){
  paste0("/media/ag-cherrmann/echernova/CTCF_anno/", name, "_CTCF_binary_filtered.csv")
})
# From imputed: only available for A549
paths2_ctcf <- "/media/ag-cherrmann/echernova/CTCF_anno/A549_CTCF_binary_imputed.csv"

paths_ctcf <- c(paths1_ctcf, paths2_ctcf)
CTCF_anno_bin <- lapply(paths_ctcf, function(path){
  return(read.csv(path, header = TRUE))
})
names(CTCF_anno_bin) <- c(cell_names_filtered, "A549")
```


# Perform integrative NMF
Additionally normalize the W matrix.

```{r}
# Ranks 2 to 10
k = 2:10

int_nmf <- run_iNMF_tensor(matrices_V, 
                           ranks = k,
                           n_initializations = 10,
                           iterations = 10^4,
                           convergence_threshold = 40,
                           extract_features = TRUE)
```
# Select optimal k
```{r}
gg_plotKStats(int_nmf)
```

#-------------------------------------------------------------------------------
# K = 3

```{r}
k=3
```

# Common H matrix
```{r}
common_h_mat <- int_nmf@HMatrix$k3
colnames(common_h_mat) <- int_nmf@input_data$colnames
rownames(common_h_mat) <- c('S_1', 'S_2', 'S_3')
h_heatmap <- Heatmap(common_h_mat,
                     col = viridis(100),
                     name = "Exposure",
                     clustering_distance_columns = 'pearson',
                     show_column_dend = TRUE,
                     show_column_names = TRUE,
                     show_row_names = TRUE,
                     cluster_rows = FALSE,
                     column_title = 'Common H matrix')
print(h_heatmap)
```


# Print individual H matrices

```{r}
list_of_h_mat <- int_nmf@HMatrix_vs$k3
lapply(names(list_of_h_mat), function(cell_name){
  
  # From named list to a matrix
  h_mat_ind <- list_of_h_mat[cell_name]
  names(h_mat_ind) <- c()
  
  #Print heatmap
  h_heatmap <- Heatmap(h_mat_ind[[1]],
                     col = viridis(100),
                     name = "Exposure",
                     clustering_distance_columns = 'pearson',
                     show_column_dend = TRUE,
                     show_column_names = TRUE,
                     show_row_names = FALSE,
                     cluster_rows = FALSE,
                     column_title = paste('H matrix: ', cell_name))
})
```
# Attribute each TAD boundary to one signature
```{r}
signatures_l <- lapply(names(list_of_h_mat), function(cell_name){
  data_matrixW <- int_nmf@WMatrix_vs$k3[cell_name]
  
  # From named list to a matrix
  names(data_matrixW) <- c()
  data_matrixW <- data.frame(data_matrixW[[1]])
  
  # Attribute maximal value to
  max_cols <- argmax(data_matrixW)
  for (row in 1:nrow(data_matrixW)){
    data_matrixW[row,] = 0
    data_matrixW[row, max_cols[row]] = 1
  }
  
  sums <- colSums(data_matrixW)
  #rownames(sums) <- c(cell_name)
  return(sums)
})
signatures <- Reduce(function(d1, d2) rbind(d1, d2), signatures_l)
colnames(signatures) <- c('S1', 'S2', 'S3')

# Calculate ratios
signatures <- signatures /rowSums(signatures)

# Melt for plotting
signatures_m <- melt(signatures)

plot <- ggplot(signatures_m, aes(x=Var2, y=value))+geom_boxplot()
print(plot)

# Mean ratios
for (n in colnames(signatures)){
  print(paste0('Mean of ', n, ': ', mean(signatures[,n])))
}
```


# Print W matrices

Conservation scores:

```{r}
lapply(names(list_of_h_mat), function(cell_name){
  data_matrixW <- int_nmf@WMatrix_vs$k3[cell_name]
  
  # From named list to a matrix
  names(data_matrixW) <- c()
  data_matrixW <- data_matrixW[[1]]
  
  # Normalize the matrix
  data_matrixW <- data_matrixW/matrixStats::rowMaxs(data_matrixW)
  
  colnames(data_matrixW) <- paste0("Sign.", 1:k)
  f1 = circlize::colorRamp2(seq(min(data_matrixW), max(data_matrixW), length = 4), c("#EEEEEE", "#FFD700", "#FF0000", "#8b0000"))
  
  # Annotation: conservation
  ann = unlist(consv_list[cell_name][[1]]['conservation'])
  col_ann = circlize::colorRamp2(seq(0, max(ann), length = 6), c('#ff0000', '#ffff00', '#66ff33', '#3399ff', '#6e34eb', '#b734eb'))
  row_ha = rowAnnotation(conservation = ann, col=list(conservation=col_ann))
  
  w_heatmap <- Heatmap(data_matrixW,
                       col = f1,
                       name = paste("W matrix: ", cell_name),
                       clustering_distance_columns = 'pearson',
                       show_column_dend = TRUE,
                       show_column_names = TRUE,
                       show_row_names = FALSE,
                       cluster_rows = TRUE,
                       cluster_columns = FALSE,
                       left_annotation = row_ha)
  w_heatmap = draw(w_heatmap)
})
```

CTCF binding scores:
```{r}
lapply(names(CTCF_anno), function(cell_name){
  data_matrixW <- int_nmf@WMatrix_vs$k3[cell_name]
  
  # From named list to a matrix
  names(data_matrixW) <- c()
  data_matrixW <- data_matrixW[[1]]
  
  # Normalize the matrix
  data_matrixW <- data_matrixW/matrixStats::rowMaxs(data_matrixW)
  
  colnames(data_matrixW) <- paste0("Sign.", 1:k)
  f1 = circlize::colorRamp2(seq(min(data_matrixW), max(data_matrixW), length = 4), c("#EEEEEE", "#FFD700", "#FF0000", "#8b0000"))
  
  # Annotation: conservation
  ann = unlist(CTCF_anno[cell_name][[1]]['CTCF'])
  col_ann = circlize::colorRamp2(seq(0, max(ann), length = 6), c('#ff0000', '#ffff00', '#66ff33', '#3399ff', '#6e34eb', '#b734eb'))
  row_ha = rowAnnotation(CTCF_binding_score = ann, col=list(CTCF_binding_score=col_ann))
  
  w_heatmap <- Heatmap(data_matrixW,
                       col = f1,
                       name = paste("W matrix: ", cell_name),
                       clustering_distance_columns = 'pearson',
                       show_column_dend = TRUE,
                       show_column_names = TRUE,
                       show_row_names = FALSE,
                       cluster_rows = TRUE,
                       cluster_columns = FALSE,
                       left_annotation = row_ha)
  w_heatmap = draw(w_heatmap)
})
```


#Attribute every TAD boundary to one signature and boxplot

Conservation
```{r}
lapply(names(list_of_h_mat), function(cell_name){
  data_matrixW <- int_nmf@WMatrix_vs$k3[cell_name][[1]]
  
  # Attribute maximal value to
  max_cols <- argmax(data_matrixW)
  for (row in 1:nrow(data_matrixW)){
    data_matrixW[row,] = 0
    data_matrixW[row, max_cols[row]] = 1
  }

  colnames(data_matrixW) <- paste0("Sign.", 1:k)

  # Add annotation
  ann = unlist(consv_list[cell_name][[1]]['conservation'])
  names(ann) <- c(1:length(ann))
  df = data.frame(signature=factor(max_cols), conservation=ann)
  
  # Boxplots
  plot <- ggplot(df, aes(x=signature, y=conservation))+
    geom_boxplot()+
    ggtitle(paste0('Distribution of TAD boundaries: ', cell_name))+
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
})
```

```{r}
#complete_df = data.frame(signature=factor(), CTCF_binding_score=numeric(), cell=character())
complete_df <- lapply(names(CTCF_anno)[1:3], function(cell_name){
  data_matrixW <- int_nmf@WMatrix_vs$k3[cell_name][[1]]
  
  # Attribute maximal value to
  max_cols <- argmax(data_matrixW)
  for (row in 1:nrow(data_matrixW)){
    data_matrixW[row,] = 0
    data_matrixW[row, max_cols[row]] = 1
  }

  colnames(data_matrixW) <- paste0("Sign.", 1:k)

  # Add annotation
  ann = unlist(CTCF_anno[cell_name][[1]]['CTCF'])
  names(ann) <- c(1:length(ann))
  df = data.frame(signature=factor(max_cols), CTCF_binding_score=ann)
  # Add data to common df
  df$cell = cell_name
  complete_df <- rbind(complete_df, df)
  
  # Student's t-test
  comparisons = compare_means(CTCF_binding_score ~ signature,  data = df, method='t.test')
  print(comparisons)
  
  return(df)
  
})

complete_df <- Reduce(function(d1, d2) rbind(d1, d2), complete_df)
complete_df$cell <- factor(complete_df$cell)
# Boxplots
plot <- ggplot(complete_df, aes(x=signature, y=CTCF_binding_score))+
        geom_violin() +
        geom_boxplot(width=0.1)+
        facet_wrap(~cell, ncol=3)+
        ggtitle('Distributions of the CTCF binding score')+
        theme(plot.title = element_text(hjust = 0.5))
png(file="/media/ag-cherrmann/echernova/figures/CTCF_distribution.png", width=550, height=300)
print(plot)
dev.off()
```
Binary CTCF
```{r}
lapply(names(CTCF_anno_bin), function(cell_name){
  data_matrixW <- int_nmf@WMatrix_vs$k3[cell_name][[1]]
  
  # Attribute maximal value to
  max_cols <- argmax(data_matrixW)
  for (row in 1:nrow(data_matrixW)){
    data_matrixW[row,] = 0
    data_matrixW[row, max_cols[row]] = 1
  }

  colnames(data_matrixW) <- paste0("Sign.", 1:3)

  # Add annotation
  ann = unlist(CTCF_anno_bin[cell_name][[1]]['CTCF'])
  names(ann) <- c(1:length(ann))
  df = data.frame(signature=factor(max_cols), CTCF_binding_score=ann)
  
  # Boxplots
  plot <- ggplot(df)+
    aes(x=signature, fill=factor(CTCF_binding_score)) +
    geom_bar(position='fill')+
    ggtitle(paste0('Distribution of TAD boundaries: ', cell_name))+
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
})
```

# Save the data
```{r}
folder_path = "/media/ag-cherrmann/echernova/integrative_nmf/"
save(int_nmf, file=paste0(folder_path, '220701_iNMF-object.RData'))
h_matrix <- int_nmf@HMatrix$k3 
colnames(h_matrix) <- int_nmf@input_data$colnames
write.csv(h_matrix, paste0(folder_path, '220701_common-H-matrix.csv'))
```





