---
title: "Signature_correlation_imputed"
output: html_document
---

```{r setup, include=FALSE}
reticulate::use_condaenv("tensorflowenv2", required = TRUE)
library(knitr)
library(circlize)
library(tidyverse)
library(ComplexHeatmap)
library(viridis)
library(BiocStyle)
library(ButchR)
library(limma)
library(ggplot2)
library(riverplot)
library(ramify)
library(ggpubr)
```

# Loading data
Prepare NMF input: loading, normalization.

```{r}
set.seed(4563)

cell_names = c("A549", "GM12878", "HepG2", "HMEC", 
             "HUVEC", "IMR90", "K562", "NHEK")

load_data <- function(name){
  
  path = paste0("/media/ag-cherrmann/echernova/model_input/", name, "/positives_", name, "_imputed.csv")
  data = read.csv(path)
  
  # Only get the central windows
  data = cbind(data[2:5], data[,grep('_0', names(data))])
  
  # Rename columns to be shorter
  new_names <- lapply(names(data[5:length(names(data))]), function(n){
    return(paste(strsplit(n, '_')[[1]][1:2], collapse = '_'))
  })
  names(data) <- c(names(data[1:4]), new_names)
  
  # Return the input matrix for NMF
  matrix_V = data.matrix(data[5:length(names(data))])

  matrix_V = drop_na(as.data.frame(matrix_V))
  
  # Normalize matrix
  matrix_V_norm <- normalizeQuantiles(matrix_V)
  
  return(matrix_V_norm)
}

matrices_V <- lapply(cell_names, load_data)

# Load TAD boundary conservation
consv_list <- lapply(cell_names, function(name){
  path_consv = paste0("/media/ag-cherrmann/echernova/tad_bound_conservation_10_40kb/", name, "_tad_bound_conservation_imputed.csv")
  consv = read.csv(path_consv)
})
```

# Perform NMF
Additionally normalize the W matrix.

```{r}
# Ranks 2 to 10
k = 2:10

all_nmf <- lapply(matrices_V, function(mat){
  
  # Correct if k is bigger than number of columns in matrix V
  if (max(k) > ncol(mat)){
    test_k = 2:(ncol(mat)-1)
  }else{
    test_k = k
  }
  
  data_nmf <- run_NMF_tensor(X = mat, 
                             ranks = test_k,
                             method = 'NMF',
                             n_initializations = 10,
                             extract_features = TRUE)
   # Normalize W matrix
  data_nmf <- normalizeW(data_nmf)
  
  return(data_nmf)
})
```

```{r}
lapply(all_nmf, function(sample_nmf){
  sample_name = strsplit(sample_nmf@input_matrix$colnames[1], '_')[[1]][1]
  print(sample_name)
  gg_plotKStats(sample_nmf)
  
})
```
Best k values:
- A549: 3, 6, 8
- GM12878: 7
- HepG2: 6
- HMEC: 3, 7
- HUVEC: 3, 6
- IMR90: 6 
- K562: 7
- NHEK: 3, 6

#-------------------------------------------------------------------------------
# K = 4
As in the non imputed data

```{r}
k=4
```


# Print H matrices

```{r}
lapply(all_nmf, function(data_nmf){
  tmp_hmatrix <- HMatrix(data_nmf, k)
  cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
  h_heatmap <- Heatmap(tmp_hmatrix,
                     col = viridis(100),
                     name = "Exposure",
                     clustering_distance_columns = 'pearson',
                     show_column_dend = TRUE,
                     show_column_names = TRUE,
                     show_row_names = FALSE,
                     cluster_rows = FALSE,
                     column_title = paste0('H matrix: ', cell_name, ', imputed'))
  # print(h_heatmap)
})
```
# Print W matrices

```{r}
for (i in c(1:length(all_nmf))){
  data_nmf = all_nmf[[i]]
  data_matrixW <- WMatrix(data_nmf, k)
  
  # Normalize the matrix
  data_matrixW <- data_matrixW/matrixStats::rowMaxs(data_matrixW)
  
  cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
  colnames(data_matrixW) <- paste0("Sign.", 1:k)
  f1 = circlize::colorRamp2(seq(min(data_matrixW), max(data_matrixW), length = 4), c("#EEEEEE", "#FFD700", "#FF0000", "#8b0000"))
  
  # Add annotation
  ann = unlist(consv_list[[i]]['conservation'])
  col_ann = circlize::colorRamp2(seq(0, max(ann), length = 6), c('#ff0000', '#ffff00', '#66ff33', '#3399ff', '#6e34eb', '#b734eb'))
  row_ha = rowAnnotation(conservation = ann, col=list(conservation=col_ann))
  
  w_heatmap <- Heatmap(data_matrixW,
                       col = f1,
                       name = paste0("W matrix: ", cell_name, ', imputed'),
                       clustering_distance_columns = 'pearson',
                       show_column_dend = TRUE,
                       show_column_names = TRUE,
                       show_row_names = FALSE,
                       cluster_rows = TRUE,
                       cluster_columns = FALSE,
                       left_annotation = row_ha)
  w_heatmap = draw(w_heatmap)
}
```
#Attribute every TAD boundary to one signature
```{r}
complete_df = data.frame(signature=factor(), conservation=numeric(), cell=character())
for (i in c(1:length(all_nmf))){
  data_nmf = all_nmf[[i]]
  data_matrixW <- WMatrix(data_nmf, k)
  
  # Attribute maximal value to
  max_cols <- argmax(data_matrixW)
  for (row in 1:nrow(data_matrixW)){
    data_matrixW[row,] = 0
    data_matrixW[row, max_cols[row]] = 1
  }
  
  cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
  colnames(data_matrixW) <- paste0("Sign.", 1:k)

  # Add annotation
  ann = unlist(consv_list[[i]]['conservation'])
  names(ann) <- c(1:length(ann))
  df = data.frame(signature=factor(max_cols), conservation=ann)
  df$cell = cell_name
  complete_df <- rbind(complete_df, df)
  
  #T-test
  comparisons = compare_means(conservation ~ signature,  data = df, method='t.test')
  print(comparisons)
  # Boxplots
  #plot <- ggplot(df, aes(x=signature, y=conservation))+
  #        geom_violin() +
  #        geom_boxplot(width=0.1)

  #print(plot)
}

plot <- ggplot(complete_df, aes(x=signature, y=conservation))+
        geom_violin() +
        geom_boxplot(width=0.1)+
        facet_wrap(~cell, ncol=4)+
        ggtitle('Distributions of conservation scores for different signatures')+
        theme(plot.title = element_text(hjust = 0.5))
png(file="/media/ag-cherrmann/echernova/figures/conservation_distribution.png", width=550, height=300)
print(plot)
dev.off()
```

# Pairwise correlation of the signatures
Pearson correlation used.

Prepare the matrix containing all signatures:

```{r}
h_mat_list <- lapply(all_nmf, function(data_nmf){
  
  # get H matrix
  h_mat <- HMatrix(data_nmf, k)
  h_mat <- data.frame(h_mat)
  
  # Add names to the rows (S#_CellName)
  row_names <- list()
  for (i in c(1:k)){
    cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
    row_names <- append(row_names, paste0('S', i, '_', cell_name))
  }
  rownames(h_mat) <- row_names
  
  # Change column names so that they are the same for all cell lines
  col_names <- lapply(names(h_mat), function(n){
    new_n <- strsplit(n, '_')[[1]][2]
    return(new_n)
  })
  names(h_mat) <- col_names
  return(h_mat)
})

# Combine all H matrices in one matrix
all_h_mat <- Reduce(function(d1, d2) rbind(d1, d2), h_mat_list)

# Transpose => columns are signatures, rows are features
all_h_mat <- t(all_h_mat)
```

Calculate Pearson correlation:
```{r}
cor_matrix <- cor(all_h_mat, method = 'pearson')
# Plot the heatmap
heat <- Heatmap(cor_matrix, name="Pearson corr.", width =unit(120, "mm"), 
                height=unit(120, "mm"),
                column_title = 'Clustering of the NMF signatures by Pearson correlation')
png(file="/media/ag-cherrmann/echernova/figures/signature_corr_imputed.png", width=550, height=500)
print(heat)
dev.off()
```




