---
title: "Signature_correlation"
output: html_document
---

```{r setup, include=FALSE}
reticulate::use_condaenv("tensorflowenv2", required = TRUE)
library(knitr)
library(circlize)
library(tidyverse)
library(ComplexHeatmap)
library(viridis)
library(BiocStyle)
library(ButchR)
library(limma)
library(ggplot2)
library(riverplot)
library(ramify)
```

# Loading data
Prepare NMF input: loading, normalization.

```{r}
set.seed(4563)

cell_names = c("K562", "GM12878", "HepG2", "IMR90")

# path_K562 = "/media/ag-cherrmann/echernova/model_input/K562/positives_K562_CTCF-from-ChIP.csv"
# path_GM12878 = "/media/ag-cherrmann/echernova/model_input/GM12878/positives_GM12878_CTCF-from-ChIP.csv"
# path_IMR90 = "/media/ag-cherrmann/echernova/model_input/IMR90/positives_IMR90.csv"
# path_HepG2 = "/media/ag-cherrmann/echernova/model_input/IMR90/positives_IMR90.csv"
# 
# paths <- c(path_K562, path_GM12878, path_IMR90)

load_data <- function(name){
  
  path = paste0("/media/ag-cherrmann/echernova/model_input/", name, "/positives_", name, "_filtered.csv")
  
  data = read.csv(path)
  
  # Only get the central windows
  data = cbind(data[2:5], data[,grep('_0', names(data))])
  
  # Rename columns to be shorter
  new_names <- lapply(names(data[5:length(names(data))]), function(n){
    return(paste(strsplit(n, '_')[[1]][1:2], collapse = '_'))
  })
  names(data) <- c(names(data[1:4]), new_names)
  
  # Return the input matrix for NMF
  matrix_V = data.matrix(data[5:length(names(data))])
  
  # Normalize matrix
  matrix_V_norm <- normalizeQuantiles(matrix_V)
  
  return(matrix_V_norm)
}

matrices_V <- lapply(cell_names, load_data)

# Load TAD boundary conservation
consv_list <- lapply(cell_names, function(name){
  path_consv = paste0("/media/ag-cherrmann/echernova/tad_bound_conservation_10_40kb/", name, "_tad_bound_conservation_filtered.csv")
  consv = read.csv(path_consv)
})
```

# Perform NMF
Additionally normalize the W matrix.

```{r}
# Ranks 2 to 10
k = 2:10

all_nmf <- lapply(matrices_V, function(mat){
  
  # Correct if k is bigger than number of columns in matrix V
  if (max(k) > ncol(mat)){
    test_k = 2:(ncol(mat)-1)
  }
  else{
    test_k = k
  }
  
  data_nmf <- run_NMF_tensor(X = mat, 
                             ranks = test_k,
                             method = 'NMF',
                             n_initializations = 10,
                             extract_features = TRUE)
  # Normalize W matrix
  data_nmf <- normalizeW(data_nmf)
  
  return(data_nmf)
})
```

```{r}
lapply(all_nmf, function(sample_nmf){
  sample_name = strsplit(sample_nmf@input_matrix$colnames[1], '_')[[1]][1]
  print(sample_name)
  gg_plotKStats(sample_nmf)
  
})
```
Best k values:
- K562: ?
- GM12878: 4
- HepG2: 8
- IMR90: 5, 8

#-------------------------------------------------------------------------------
# K = 4

```{r}
k=4
```


# Print H matrices

```{r}
lapply(all_nmf, function(data_nmf){
  tmp_hmatrix <- HMatrix(data_nmf, k)
  cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
  h_heatmap <- Heatmap(tmp_hmatrix,
                     col = viridis(100),
                     name = "Exposure",
                     clustering_distance_columns = 'pearson',
                     show_column_dend = TRUE,
                     show_column_names = TRUE,
                     show_row_names = FALSE,
                     cluster_rows = FALSE,
                     column_title = paste('H matrix: ', cell_name))
  # print(h_heatmap)
})
```
# Print W matrices

```{r}
for (i in c(1:length(all_nmf))){
  data_nmf = all_nmf[[i]]
  data_matrixW <- WMatrix(data_nmf, k)
  
  # Normalize the matrix
  data_matrixW <- data_matrixW/matrixStats::rowMaxs(data_matrixW)
  
  cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
  colnames(data_matrixW) <- paste0("Sign.", 1:k)
  f1 = circlize::colorRamp2(seq(min(data_matrixW), max(data_matrixW), length = 4), c("#EEEEEE", "#FFD700", "#FF0000", "#8b0000"))
  
  ann = unlist(consv_list[[i]]['conservation'])
  col_ann = circlize::colorRamp2(seq(0, max(ann), length = 6), c('#ff0000', '#ffff00', '#66ff33', '#3399ff', '#6e34eb', '#b734eb'))
  row_ha = rowAnnotation(conservation = ann, col=list(conservation=col_ann))
  
  w_heatmap <- Heatmap(data_matrixW,
                       col = f1,
                       name = paste("W matrix: ", cell_name),
                       clustering_distance_columns = 'pearson',
                       show_column_dend = TRUE,
                       show_column_names = TRUE,
                       show_row_names = FALSE,
                       cluster_rows = TRUE,
                       cluster_columns = FALSE,
                       left_annotation = row_ha)
  w_heatmap = draw(w_heatmap)
}
```
#Attribute every TAD boundary to one signature and boxplot
```{r}
for (i in c(1:length(all_nmf))){
  data_nmf = all_nmf[[i]]
  data_matrixW <- WMatrix(data_nmf, k)
  
  # Attribute maximal value to
  data_matrixW <- WMatrix(data_nmf, k)
  max_cols <- argmax(data_matrixW)
  for (row in 1:nrow(data_matrixW)){
    data_matrixW[row,] = 0
    data_matrixW[row, max_cols[row]] = 1
  }
  
  cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
  colnames(data_matrixW) <- paste0("Sign.", 1:k)

  # Add annotation
  ann = unlist(consv_list[[i]]['conservation'])
  names(ann) <- c(1:length(ann))
  df = data.frame(signature=factor(max_cols), conservation=ann)
  
  # Boxplots
  plot <- ggplot(df, aes(x=signature, y=conservation))+
    geom_boxplot()+
    ggtitle(paste0('Distribution of TAD boundaries: ', cell_name))+
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
}
```

# Pairwise correlation of the signatures
Pearson correlation used.

Prepare the matrix containing all signatures:

```{r}
# Some cells have more input data than other -> eliminate
elim_names <- c('H3K27me3', 'H3K17me3', 'H3K9me1')

h_mat_list <- lapply(all_nmf, function(data_nmf){
  
  # get H matrix
  h_mat <- HMatrix(data_nmf, k)
  h_mat <- data.frame(h_mat)
  
  # eliminate columns
  for (elim_n in elim_names){
    if (length(grep(elim_n, names(h_mat)))>0){
    h_mat <- h_mat[, -grep(elim_n, names(h_mat))]
    }
  }
  
  # Add names to the rows (S#_CellName)
  row_names <- list()
  for (i in c(1:k)){
    cell_name <- strsplit(data_nmf@input_matrix$colnames[1], '_')[[1]][1]
    row_names <- append(row_names, paste0('S', i, '_', cell_name))
  }
  rownames(h_mat) <- row_names
  
  # Change column names so that they are the same for all cell lines
  col_names <- lapply(names(h_mat), function(n){
    new_n <- strsplit(n, '_')[[1]][2]
    return(new_n)
  })
  names(h_mat) <- col_names
  return(h_mat)
})
#K562 has a wrong name of one column
# names(h_mat_list[[1]])[2] <- "H2AFZ"

# Combine all H matrices in one matrix
all_h_mat <- Reduce(function(d1, d2) rbind(d1, d2), h_mat_list)

# Transpose => columns are signatures, rows are features
all_h_mat <- t(all_h_mat)
```

Calculate Pearson correlation:
```{r}
cor_matrix <- cor(all_h_mat, method = 'pearson')
# Plot the heatmap
heat <- Heatmap(cor_matrix, name="Pearson corr.", width =unit(80, "mm"), 
                height=unit(80, "mm"), 
                column_title = 'Clustering of the NMF signatures by Pearson correlation')

png(file="/media/ag-cherrmann/echernova/figures/signature_corr.png")
print(heat)
dev.off()
```




